---
title: "variable_breakdown_1"
author: "dawon"
date: "2025-06-26"
output: html_document
---

#import packages
```{python setup, include=FALSE} 
import pandas as pd
import os
import seaborn as sns
import matplotlib.pyplot as plt
```


```{python, include=FALSE}
# Define file paths
folder_path = r"C:\Users\ohdaw\OneDrive - Harvard University\tu_data"
file_path = os.path.join(folder_path, "data_reduced_hh_0to9_no_edu1.csv")
#output_path = os.path.join(folder_path, ".csv")

# Load dataset
dt = pd.read_csv(file_path, encoding="utf-8-sig")
```

```{python}
counts = dt.groupby('PrimTargetPurp_Cat')['JourneyId'].nunique().sort_index()
percentages = dt['PrimTargetPurp_Cat'].value_counts(normalize=True).sort_index() * 100

purpose_summary = pd.DataFrame({
    'PrimTargetPurp_Cat': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

purpose_summary['Percentage'] = purpose_summary['Percentage'].round(2)

purpose_summary['Purpose'] = purpose_summary['PrimTargetPurp_Cat']

purpose_summary = purpose_summary[['Purpose', 'Count', 'Percentage']]

print(purpose_summary)
```

```{python}
dt_before = dt.copy()

# Check original length
initial_rows = len(dt)

# Drop NAs step by step and count
dt = dt.dropna(subset=['inc_quant'])
after_inc = len(dt)
dropped_inc = initial_rows - after_inc

dt = dt.dropna(subset=['HousehNumAdults'])
after_adults = len(dt)
dropped_adults = after_inc - after_adults

dt = dt.dropna(subset=['IsWorker'])
after_worker = len(dt)
dropped_worker = after_adults - after_worker

dt = dt.dropna(subset=['OlderInCouple'])
after_couple = len(dt)
dropped_couple = after_worker - after_couple

total_dropped = initial_rows - after_couple

print("Dropped due to inc_quant:", dropped_inc)
print("Dropped due to HousehNumAdults:", dropped_adults)
print("Dropped due to IsWorker:", dropped_worker)
print("Dropped due to OlderInCouple:", dropped_couple)
print("Total dropped:", total_dropped)
print("Remaining rows:", after_couple)

# Count unique IDs after cleaning
unique_journeys = dt["JourneyId"].nunique()
unique_sessions = dt["SessionId"].nunique()

print(f"Unique JourneyId: {unique_journeys}")
print(f"Unique SessionId: {unique_sessions}")

# Before dropping NAs
initial_journeys = dt_before["JourneyId"].nunique()
initial_sessions = dt_before["SessionId"].nunique()

# After dropping NAs (dt is your cleaned dataset)
final_journeys = dt["JourneyId"].nunique()
final_sessions = dt["SessionId"].nunique()

# Differences
dropped_journeys = initial_journeys - final_journeys
dropped_sessions = initial_sessions - final_sessions

print("Unique JourneyId before:", initial_journeys)
print("Unique JourneyId after:", final_journeys)
print("Dropped JourneyId:", dropped_journeys)

print("Unique SessionId before:", initial_sessions)
print("Unique SessionId after:", final_sessions)
print("Dropped SessionId:", dropped_sessions)
```
```{python}
# Count unique SessionId by inc_quant
session_counts = (
    dt.groupby("inc_quant")["SessionId"]
       .nunique()
       .reset_index(name="UniqueSessions")
)

# Add percentage column
total_sessions = session_counts["UniqueSessions"].sum()
session_counts["Percentage"] = (session_counts["UniqueSessions"] / total_sessions * 100).round(2)

print(session_counts)
```


# Percentage of each purposes (1)
```{python}
counts = dt.groupby('PrimTargetPurp_Cat')['JourneyId'].nunique().sort_index()
percentages = dt['PrimTargetPurp_Cat'].value_counts(normalize=True).sort_index() * 100

purpose_summary = pd.DataFrame({
    'PrimTargetPurp_Cat': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

purpose_summary['Percentage'] = purpose_summary['Percentage'].round(2)

purpose_summary['Purpose'] = purpose_summary['PrimTargetPurp_Cat']

purpose_summary = purpose_summary[['Purpose', 'Count', 'Percentage']]

print(purpose_summary)
```

## Purpose categories
1: "home residence"
11, 50, 12, 13, 14: "work and education"
20, 23, 31, 32, 33: "errands and shopping"
21, 22: "chauffeuring"
25, 38, 42, 43, 45, 47, 49: "leisure and social activities"


```{python}
# Step 1: Get total travel time per JourneyId
duration_by_journey = dt.groupby("JourneyId")["TravelTimeMinutes"].sum().reset_index()

# Step 2: Merge with purpose (PrimTargetPurp_Cat)
# Assuming each JourneyId has one PrimTargetPurp_Cat value — adjust if needed
journey_purpose = dt[["JourneyId", "PrimTargetPurp_Cat"]].drop_duplicates()

duration_with_purpose = duration_by_journey.merge(journey_purpose, on="JourneyId", how="left")

# Step 3: Get list of unique purposes
categories = duration_with_purpose["PrimTargetPurp_Cat"].dropna().unique()

# Step 4: Plot histogram for each category
for cat in categories:
    subset = duration_with_purpose[duration_with_purpose["PrimTargetPurp_Cat"] == cat]

    plt.figure(figsize=(8, 4))
    plt.hist(subset["TravelTimeMinutes"], bins=72, range=(0, 720), color='skyblue', edgecolor='black')
    plt.title(f'Distribution of Journey Duration for "{cat}"')
    plt.xlabel('Journey Duration (minutes)')
    plt.ylabel('Count')
    plt.grid(True, linestyle='--', alpha=0.5)
    plt.tight_layout()
    plt.show()
```

# Distance/ travel time of each purposes (1)
```{python}
purpose_stats = dt.groupby("PrimTargetPurp_Cat").agg(
    Count=("JourneyId", "nunique"),
    AvgDistance_km=("TransportDistance", "mean"),
    MedianDistance_km=("TransportDistance", "median"),
    MinDistance_km=("TransportDistance", "min"),
    MaxDistance_km=("TransportDistance", "max"),
    StdDistance_km=("TransportDistance", "std"),
    AvgTravelTime_min=("TravelTimeMinutes", "mean"),
    MedianTravelTime_min=("TravelTimeMinutes", "median"),
    MinTravelTime_min=("TravelTimeMinutes", "min"),
    MaxTravelTime_min=("TravelTimeMinutes", "max"),
    StdTravelTime_min=("TravelTimeMinutes", "std")
).reset_index()

purpose_stats["Purpose"] = purpose_stats["PrimTargetPurp_Cat"]

purpose_stats = purpose_stats[
    ["Purpose", "Count", "AvgDistance_km", "MedianDistance_km", "MinDistance_km", "MaxDistance_km", "StdDistance_km", "AvgTravelTime_min", "MedianTravelTime_min", "MinTravelTime_min", "MaxTravelTime_min", "StdTravelTime_min"]
]

purpose_stats = purpose_stats.round(2)
```
```{r}
library(dplyr)
library(knitr)
library(reticulate)

reticulate::py$purpose_stats |>
  arrange(desc(Count)) |>
  kable(digits = 2, caption = "Trip Purpose Summary Statistics")
```


# Percentage of each modes (1)
```{python}
counts = dt.groupby('primode_cat')['JourneyId'].nunique().sort_index()
percentages = dt['primode_cat'].value_counts(normalize=True).sort_index() * 100

mode_summary = pd.DataFrame({
    'primode_cat': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

mode_summary['Percentage'] = mode_summary['Percentage'].round(2)

mode_summary['Mode'] = mode_summary['primode_cat']

mode_summary = mode_summary[['Mode', 'Count', 'Percentage']]

print(mode_summary)
```

## Mode categories
1: "Walk" 
2: "Bicycle"
11, 12, 13, 14: "Motorized Transport"
25, 31, 32, 33, 34, 35, 37, 41: "Public Transport"

# Percentage of each modes (2)
```{python}
mode_map = {
    1: "Walk or run",
    2: "Bicycle",
    3: "Moped 30",
    4: "Moped 45",
    5: "Skateboard/roller skates/scooter",
    6: "Horse carriage, horse",
    7: "Disability moped (electric)",
    8: "Electric scooter etc.",
    11: "Passenger car",
    12: "Van",
    13: "Lorry",
    14: "Motorcycle",
    15: "Tractor, working vehicle",
    25: "Taxi cab",
    26: "Tourist coach, rented bus",
    31: "Collective, Public bus",
    32: "S-train",
    33: "Other train",
    34: "Metro train",
    35: "Dial-a-ride, flexible transport service",
    37: "Light rail/tram",
    41: "Ferry, water bus",
    42: "Pleasure boat",
    51: "Airplane"
}

counts = dt.groupby('PrimMode')['JourneyId'].nunique().sort_index()
percentages = dt['PrimMode'].value_counts(normalize=True).sort_index() * 100

mode_summary = pd.DataFrame({
    'PrimMode': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

mode_summary['Percentage'] = mode_summary['Percentage'].round(2)

mode_summary['Mode'] = mode_summary['PrimMode'].map(mode_map)

mode_summary = mode_summary[['Mode', 'Count', 'Percentage']]

print(mode_summary)
```

# Distance/ travel time of each modes (1)
```{python}
purpose_stats = dt.groupby("primode_cat").agg(
    Count=("JourneyId", "nunique"),
    AvgDistance_km=("SumLen", "mean"),
    MedianDistance_km=("SumLen", "median"),
    MinDistance_km=("SumLen", "min"),
    MaxDistance_km=("SumLen", "max"),
    StdDistance_km=("SumLen", "std"),
    AvgTravelTime_min=("SumMin", "mean"),
    MedianTravelTime_min=("SumMin", "median"),
    MinTravelTime_min=("SumMin", "min"),
    MaxTravelTime_min=("SumMin", "max"),
    StdTravelTime_min=("SumMin", "std")
).reset_index()

purpose_stats["Mode"] = purpose_stats["primode_cat"]

purpose_stats = purpose_stats[
    ["Mode", "Count", "AvgDistance_km", "MedianDistance_km", "MinDistance_km", "MaxDistance_km", "StdDistance_km",
     "AvgTravelTime_min", "MedianTravelTime_min", "MinTravelTime_min", "MaxTravelTime_min", "StdTravelTime_min"]
]

purpose_stats = purpose_stats.round(2)
```
```{r}
reticulate::py$purpose_stats |>
  arrange(desc(Count)) |>
  kable(digits = 2, caption = "Mode Summary Statistics")
```

# Distance/ travel time of each modes (2)
```{python}
purpose_stats = dt.groupby("PrimMode").agg(
    Count=("JourneyId", "nunique"),
    AvgDistance_km=("SumLen", "mean"),
    MedianDistance_km=("SumLen", "median"),
    MinDistance_km=("SumLen", "min"),
    MaxDistance_km=("SumLen", "max"),
    StdDistance_km=("SumLen", "std"),
    AvgTravelTime_min=("SumMin", "mean"),
    MedianTravelTime_min=("SumMin", "median"),
    MinTravelTime_min=("SumMin", "min"),
    MaxTravelTime_min=("SumMin", "max"),
    StdTravelTime_min=("SumMin", "std")
).reset_index()

purpose_stats["Mode"] = purpose_stats["PrimMode"].map(mode_map) 

purpose_stats = purpose_stats[
    ["Mode", "Count", "AvgDistance_km", "MedianDistance_km", "MinDistance_km", "MaxDistance_km", "StdDistance_km",
     "AvgTravelTime_min", "MedianTravelTime_min", "MinTravelTime_min", "MaxTravelTime_min", "StdTravelTime_min"]
]

purpose_stats = purpose_stats.round(2)
```
```{r}
reticulate::py$purpose_stats |>
  arrange(desc(Count)) |>
  kable(digits = 2, caption = "Mode Summary Statistics")
```

#  Number of Journey Types with children 0-10 (maybe 10-17 or not)
```{python}
counts = dt.groupby('JourneyWithChildren')['JourneyId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Journey Type  with children 0-17': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```

#  Number of accompanied children on a journey 0-9
```{python}
counts = dt.groupby('PartyNum017_max')['JourneyId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Number of accompanied children on a journey 0-9': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```
## Number of accompanied children on a journey 0-9
```{python}
counts = dt.groupby('PartyNumu10_max')['JourneyId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Number of accompanied children on a journey under 10y': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```
#  Number of accompanied children on a journey 10-17
```{python}
counts = dt.groupby('PartyNum1017_max')['JourneyId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Number of accompanied children on a journey 10-17': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```

#  Number of children under 6y in a household
```{python}
counts = dt.groupby('Chldu6')['SessionId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Number of children under 6y in a household': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```

```{python}
dt["Chldu6_fam"] = dt["FamNumPers"] - dt["FamNumPersO6"]

counts = dt.groupby('Chldu6_fam')['SessionId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Number of children under 6y in a family': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```

```{python}
counts = dt.groupby('NumChld')['SessionId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Number of children under 18y in a household': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```

```{python}
dt["NumChld_fam"] = dt["FamNumPers"] - dt["FamNumAdults"]

counts = dt.groupby('NumChld_fam')['SessionId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Number of children under 18y in a family': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```

# Journey complexity
```{python}
counts = dt.groupby('JourneyComplx')['JourneyId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Journey complexity': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```

# Gender of the caregivers (1: Male/ 2: Female)
```{python}
counts = dt.groupby('RespSex')['JourneyId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Gender': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```

# Work status of the caregivers (1: Worker/ 0: Non-worker)
```{python}
counts = dt.groupby('IsWorker')['JourneyId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Worker': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```

#Number of children (Age 0-17) in a HH
```{python}
counts = dt.groupby('Num_0_17')['SessionId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Number ofchildren(0-17) in a Family': counts.index,
    'Count': counts.values,
    'Percentage': percentages.round(2)
})
print(summary)
```

#Number of younger Sibling(Age 0-9) in a HH
```{python}
counts = dt.groupby('Num_0_9')['SessionId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Number of younger Siblings (Age 0–9) in a Family': counts.index,
    'Count': counts.values,
    'Percentage': percentages.round(2)
})
print(summary)
```

#Number of Older Sibling(Age 10-17) in a HH
```{python}
counts = dt.groupby('Num_10_17')['SessionId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Number of Older Siblings (Age 10–17) in a Family': counts.index,
    'Count': counts.values,
    'Percentage': percentages.round(2)
})
print(summary)
```

# Number of adults in a HH
```{python}
counts = dt.groupby('FamNumAdults')['SessionId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Number of adults in a Family': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```

```{python}
counts = dt.groupby('HousehNumAdults')['SessionId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Number of adults in a household': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```

# Position in the nuclear family
```{python}
family_map = {
    10: "Single",
    11: "Older in couple",
    12: "Younger in couple",
    20: "Child in nuclear family"
}

dt['FamilyPositionLabel'] = dt['PosInFamily'].map(family_map)

dt['FamilyPositionLabel'] = dt['FamilyPositionLabel'].fillna('Missing')

counts = dt.groupby('FamilyPositionLabel')['SessionId'].nunique()
percentages = counts / counts.sum() * 100

summary = pd.DataFrame({
    'Position in the nuclear family': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

summary['Percentage'] = summary['Percentage'].round(2)

print(summary)
```
## 97 missing JourneyIds

# But how many adults in the missing category?
```{python}
# Filter rows where PosInFamily is missing
missing_family = dt[dt['PosInFamily'].isna()]

# Count unique journeys by number of adults in the family for missing PosInFamily
counts = missing_family.groupby('FamNumAdults')['SessionId'].nunique()
percentages = counts / counts.sum() * 100

# Create summary DataFrame
summary_missing = pd.DataFrame({
    'Number of adults in household': counts.index,
    'Count': counts.values,
    'Percentage': percentages.values
})

# Round and sort
summary_missing['Percentage'] = summary_missing['Percentage'].round(2)
summary_missing = summary_missing.sort_values('Number of adults in household')

print(summary_missing)
```

# Total Distance Traveled
```{python}
distance_by_journey = dt.groupby('JourneyId')['TransportDistance'].sum()

# Plot histogram with 300 bins in the range 0–2000 km
plt.figure(figsize=(8, 4))
plt.hist(distance_by_journey, bins=2000, range=(0, 200), color='skyblue', edgecolor='black')
plt.xlabel('Total Distance per Journey (km)')
plt.ylabel('Count')
plt.title('Distribution of Transport Distances (0–200 km)')
plt.grid(True, linestyle='--', alpha=0.5)
plt.tight_layout()
plt.show()
```


```{python}
distance_by_journey = dt.groupby('JourneyId')['TransportDistance'].sum()

# Plot histogram with 300 bins in the range 0–2000 km
plt.figure(figsize=(8, 4))
plt.hist(distance_by_journey, bins=8500, range=(0, 850), color='skyblue', edgecolor='black')
plt.xlabel('Total Distance per Journey (km)')
plt.ylabel('Count')
plt.title('Distribution of Transport Distances (0–850 km)')
plt.grid(True, linestyle='--', alpha=0.5)
plt.tight_layout()
plt.show()
```


```{python}
distance_by_journey_max = dt.groupby('JourneyId')['TransportDistance'].max()
max_distance = distance_by_journey_max.max()

print(f"Maximum Distance Traveled: {max_distance:.2f}")
```

# Total Journey Time
```{python}
distance_by_journey = dt.groupby('JourneyId')['TravelTimeMinutes'].sum()

# Plot histogram with 300 bins in the range 0–2000 km
plt.figure(figsize=(8, 4))
plt.hist(distance_by_journey, bins=72, range=(0, 720), color='skyblue', edgecolor='black')
plt.xlabel('Journeys Duration (minutes)')
plt.ylabel('Count')
plt.title('Distribution of Journeys Duration Under 12h')
plt.grid(True, linestyle='--', alpha=0.5)
plt.tight_layout()
plt.show()
```

```{python}
percent_over_720 = (dt['TravelTimeMinutes'] > 720).mean() * 100
print(f"Percent of travel times over 720 minutes: {percent_over_720:.2f}%")
```

# Total Transport Time
```{python}
distance_by_journey = dt.groupby('JourneyId')['SumMin'].sum()

# Plot histogram with 300 bins in the range 0–2000 km
plt.figure(figsize=(8, 4))
plt.hist(distance_by_journey, bins=70, range=(0, 350), color='skyblue', edgecolor='black')
plt.xlabel('Total Time per Journey (minutes)')
plt.ylabel('Count')
plt.title('Distribution of Transport Time')
plt.grid(True, linestyle='--', alpha=0.5)
plt.tight_layout()
plt.show()
```

#Total Activity Time
```{python}
distance_by_journey = dt.groupby('JourneyId')['ActivityTime'].sum()

# Plot histogram with 300 bins in the range 0–2000 km
plt.figure(figsize=(8, 4))
plt.hist(distance_by_journey, bins=72, range=(0, 720), color='skyblue', edgecolor='black')
plt.xlabel('Total Time per Journey (minutes)')
plt.ylabel('Count')
plt.title('Distribution of Activity Time')
plt.grid(True, linestyle='--', alpha=0.5)
plt.tight_layout()
plt.show()
```

```{python}
counts = dt.groupby(['PrimTargetPurp_Cat', 'PartyNumu10_max']).size().reset_index(name='Count')

print(counts)
```

```{python}
counts = dt.groupby(['PrimTargetPurp_Cat', 'JourneyWithChildren']).size().reset_index(name='Count')

print(counts)
```

```{python}
journey_df = dt.groupby(['JourneyId', 'PrimTargetPurp_Cat'])['TravelTimeMinutes'].mean().reset_index()

for purpose in journey_df['PrimTargetPurp_Cat'].unique():
    subset = journey_df[journey_df['PrimTargetPurp_Cat'] == purpose]

    plt.figure(figsize=(8, 4))
    sns.histplot(subset['TravelTimeMinutes'], bins=30, kde=True)
    plt.title(f"Distribution of Journey Duration for Each Purpose: {purpose}")
    plt.xlabel("Journey Duration (minutes)")
    plt.ylabel("Count")
    plt.tight_layout()
    plt.show()
```

```{python}
counts = dt.groupby(['PrimTargetPurp_Cat', 'SingleAdult']).size().reset_index(name='Count')

print(counts)
```

```{python}
counts = dt.groupby(['PrimTargetPurp_Cat', 'HousehNumAdults']).size().reset_index(name='Count')

print(counts)
```

```{python}
journey_df = dt.groupby(['JourneyId', 'PrimTargetPurp_Cat'])['RespAgeSimple'].mean().reset_index()

for age in journey_df['PrimTargetPurp_Cat'].unique():
    subset = journey_df[journey_df['PrimTargetPurp_Cat'] == age]

    plt.figure(figsize=(8, 4))
    sns.histplot(subset['RespAgeSimple'], bins=30, kde=True)
    plt.title(f"Distribution of Caregiver's Age for Each Purpose: {age}")
    plt.xlabel("Caregiver's Agee (years)")
    plt.ylabel("Count")
    plt.tight_layout()
    plt.show()
```

```{python}
counts = (
    dt.groupby(['PosInFamily', 'RespSex'])['SessionId']
    .nunique()
    .reset_index(name='Count')
)

print(counts)

# 10	Single	
# 11	Older in couple	
# 12	Younger in couple
```
```{python}
purpose_map = {
    11: "Workplace",
    12: "School, educational institution",
    13: "Youth center, youth club, after-school center",
    14: "Nursery, creche, day care",
    20: "(Unknown Errand)",
    21: "Escorting to/from activity",
    22: "Escorting to/from transport",
    23: "Collect/bring objects",
    25: "(Unknown leisure)",
    31: "Shopping",
    32: "Other errand",
    33: "Social/health",
    38: "Church, Religious services",
    42: "Do sports",
    43: "Entertainment",
    45: "Leisure round trip",
    47: "Meetings in private context",
    49: "Other leisure activity"
}

journey_counts = dt.groupby('PrimTargetPurp')['JourneyId'].nunique().sort_index()

journey_percentages = (journey_counts / journey_counts.sum() * 100).round(2)

purpose_summary = pd.DataFrame({
    'PrimTargetPurp': journey_counts.index,
    'Count': journey_counts.values,
    'Percentage': journey_percentages.values,
    'Purpose': journey_counts.index.map(purpose_map)
})

purpose_child_counts = (
    dt.groupby(['PrimTargetPurp', 'JourneyWithChildren'])['JourneyId']
    .nunique()
    .reset_index(name='Count')
)

print("=== Purpose Summary ===")
print(purpose_summary)

print("\n=== Journey Counts by Purpose & JourneyWithChildren ===")
print(purpose_child_counts)
```

```{python}
cols = [
    'JourneyWithChildren',
    'primode_cat',
    'PrimTargetPurp_Cat',
    'inc_quant',
    'ChildinHH_1017',
    'YoungerInCouple',
    'IsWoman',
    'IsWorker',
    'JourneyComplx',
    'RespSex'
]

one_hot = pd.get_dummies(dt[cols], columns=cols[1:])

summary = one_hot.groupby('JourneyWithChildren').sum()

with pd.option_context('display.max_rows', None, 'display.max_columns', None):
    print(summary)
```

```{python}
cols = [
    'JourneyWithChildren',
    'TravelTimeMinutes',
    'IncHouseh',
    'HousehNumPers',
    'HousehNumAdults',
    'ChildinHH_09',
    'Chldu6',
    'PartyNumu10_max',
    'RespAgeSimple'
]

summary_stats = dt[cols].groupby('JourneyWithChildren').agg(['mean', 'median', 'min', 'max', 'std'])

with pd.option_context('display.max_rows', None, 'display.max_columns', None):
    print(summary_stats)
```

```{python}
# Step 1: Count number of unique JourneyId's per SessionId
journeys_per_session = dt.groupby('SessionId')['JourneyId'].nunique()

# Step 2: Count how many SessionIds fall into each journey count group
summary = journeys_per_session.value_counts().sort_index().reset_index()
summary.columns = ['JourneyId per SessionId', 'Count']

# Step 3: Print the table
print(summary)
```



#Filter out 'missing' cat in 'PosInFamily'
```{python}
family_map = {
    10: "Single",
    11: "Older in couple",
    12: "Younger in couple",
    20: "Child in nuclear family"
}

dt['FamilyPositionLabel'] = dt['PosInFamily'].map(family_map)
dt['FamilyPositionLabel'] = dt['FamilyPositionLabel'].fillna('Missing')

# 2. Filter out rows with 'Missing' label
filtered_dt = dt[dt['FamilyPositionLabel'] != 'Missing'].copy()

# 3. Save to CSV
folder_path = r"C:\Users\ohdaw\OneDrive - Harvard University\tu_data"
output_file = os.path.join(folder_path, "data_reduced_hh_0to9_filtered.csv")
filtered_dt.to_csv(output_file, index=False, encoding="utf-8-sig")
```

##Additional: Filter out Education
```{python}
folder_path = r"C:\Users\ohdaw\OneDrive - Harvard University\tu_data"
input_file = os.path.join(folder_path, "data_reduced_hh_0to9_filtered.csv")
df = pd.read_csv(input_file, encoding="utf-8-sig")

# 2. Filter out 'education' in 'PrimTargetPurp_Cat'
df_filtered = df[df['PrimTargetPurp_Cat'].str.lower() != 'education'].copy()

# 3. Save the filtered data
output_file = os.path.join(folder_path, "data_reduced_hh_0to9_filtered_no_edu.csv")
df_filtered.to_csv(output_file, index=False, encoding="utf-8-sig")
```

```{python}
df_filtered = df_filtered.dropna(subset=['inc_quant'])
df_filtered = df_filtered.dropna(subset=['HousehNumAdults'])
df_filtered = df_filtered.dropna(subset=['IsWorker'])
df_filtered = df_filtered.dropna(subset=['OlderInCouple'])
```

```{python}
worker_counts = df_filtered['IsWorker'].value_counts()

worker_percentages = df_filtered['IsWorker'].value_counts(normalize=True) * 100

worker_summary = pd.DataFrame({
    'Count': worker_counts,
    'Percentage': worker_percentages
})

print(worker_summary)
```
```{python}
older_counts = df_filtered['OlderInCouple'].value_counts()

older_percentages = df_filtered['OlderInCouple'].value_counts(normalize=True) * 100

older_summary = pd.DataFrame({
    'Count': older_counts,
    'Percentage': older_percentages
})

print(older_summary)
```
```{python}
women_counts = df_filtered['IsWoman'].value_counts()

women_percentages = df_filtered['IsWoman'].value_counts(normalize=True) * 100

women_summary = pd.DataFrame({
    'Count': women_counts,
    'Percentage': women_percentages
})

print(older_summary)
```
```{python}
cmplx_counts = df_filtered['JourneyComplx'].value_counts()

cmplx_percentages = df_filtered['JourneyComplx'].value_counts(normalize=True) * 100

cmplx_summary = pd.DataFrame({
    'Count': cmplx_counts,
    'Percentage': cmplx_percentages
})

print(cmplx_summary)
```
```{python}
purp_counts = df_filtered['PrimTargetPurp_Cat'].value_counts()

purp_percentages = df_filtered['PrimTargetPurp_Cat'].value_counts(normalize=True) * 100

purp_summary = pd.DataFrame({
    'Count': purp_counts,
    'Percentage': purp_percentages
})

print(purp_summary)
```
```{python}
mode_counts = df_filtered['primode_cat'].value_counts()

mode_percentages = df_filtered['primode_cat'].value_counts(normalize=True) * 100

mode_summary = pd.DataFrame({
    'Count': mode_counts,
    'Percentage': mode_percentages
})

print(mode_summary)
```



#Non-workers in purpose_work model
```{python}
df_filtered_work = df_filtered[df_filtered['PrimTargetPurp_Cat'] == 'work']

worker_counts = df_filtered_work['IsWorker'].value_counts()

worker_percentages = df_filtered_work['IsWorker'].value_counts(normalize=True) * 100

worker_summary = pd.DataFrame({
    'Count': worker_counts,
    'Percentage': worker_percentages
})

print(worker_summary)
```



```{python}
import pandas as pd
import numpy as np

# Apply log transformation (replace 0 or negative with NaN to avoid math error)
df_filtered['Log_TravelTimeMinutes'] = np.log(df_filtered['TravelTimeMinutes'].replace(0, np.nan))

# Variables to summarize
variables = ['Log_TravelTimeMinutes', 'Chldu6', 'Num_10_17', 'HousehNumAdults']

# Compute min, max, mean, std
summary_stats = df_filtered[variables].agg(['min', 'max', 'mean', 'std']).transpose()

# Rename and format output
summary_stats.index.name = 'Variable'
summary_stats.reset_index(inplace=True)

print(summary_stats)

```

```{python}
# List of variables to check
variables = ['TravelTimeMinutes', 'Chldu6', 'Num_10_17', 'HousehNumAdults']

# Count non-NaN values
non_nan_counts = df_filtered[variables].count()

# Convert to DataFrame for clearer presentation (optional)
non_nan_counts_df = non_nan_counts.reset_index()
non_nan_counts_df.columns = ['Variable', 'Non-NaN Count']

print(non_nan_counts_df)

```
```{python}
# Count of each category
count_series = df_filtered['inc_quant'].value_counts(dropna=False).sort_index()

# Percentage of each category
percentage_series = df_filtered['inc_quant'].value_counts(normalize=True, dropna=False).sort_index() * 100

# Combine into a DataFrame
income_summary = pd.DataFrame({
    'Count': count_series,
    'Percentage': percentage_series.round(3)
}).reset_index()

income_summary.columns = ['Income Quintile', 'Count', 'Percentage (%)']

print(income_summary)

```

